---
title: "Palettes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Palettes

Below is a collection of palettes for use in the app or in reports.

```{r}
# Initilize libraries
library(ggplot2)
# Utility function to display palette
library(scales)

#Palettes
library(RColorBrewer)
library(viridisLite)
library(ggsci)
library(Polychrome)
library(Seurat)
```

## Categorical Palettes

### Seurat Palletes

Categorical palettes used by Seurat are accessed via `Seurat::DiscretePalette()`. These palettes are designed for large numbers of colors.

```{r}
# Loop through all Seurat palettes
palettes <- list("alphabet", "alphabet2", "glasbey", "stepped", "polychrome")

for (palette_name in palettes){
  show_col(DiscretePalette(20, palette = palette_name))
}
```


### Palletes from ggsci package

The [ggsci package](https://nanx.me/ggsci/articles/ggsci.html) contains several categorical palettes inspired by scientific journals and science fiction. The palettes usually contain 7-10 colors, and they can be extended with `colorRampPalette()`.

```{r}
# AAAS colors: only 10 present
show_col(pal_aaas("default")(10))

# Extend to 20 colors
show_col(
  colorRampPalette(
    # Original 10 colors (pal_aaas("default") returns a function, to which
    # n is passed)
    pal_aaas("default")(10)
    )(20)
  )

# This is a deterministic approach (same results are returned each time the function is ran)
```

```{r}
# Nature Publishing Group Palette: Original 10 colors
show_col(pal_npg("nrc")(10))
```


```{r}
# Extend Nature Publishing Group Palette
show_col(
  colorRampPalette(
    # Original 10 colors 
    pal_npg("nrc")(10)
    # Call '20' from function produced to get 20 colors
    )(20)
  )
```

```{r}
# Extend Nature Publishing Group Palette
show_col(
  colorRampPalette(
    # Original 10 colors 
    pal_npg("nrc")(10)
    # Call '20' from function produced to get 20 colors
    )(6)
  )
```


```{r}
# NJEM Pallete
# Original Colors
original <- pal_nejm()(8)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))
```

```{r}
# Lancet Pallete
# Original Colors
original <- pal_lancet()(9)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))
```

```{r}
# JAMA palette
# Original Colors
original <- pal_jama()(7)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))
```

```{r}
# Journal of Clinical Oncology Pallette
# Original Colors
original <- pal_jco()(10)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))
```

```{r}
# UCSC Genome Browser Palette
# Original palette has 26 colors; does not need extending
show_col(pal_ucscgb()(26))
```

```{r}
# D3.js Palette
# D3 is a JavaScript library for interactive data 
# visualization (https://d3js.org/)
# Original Colors
original <- pal_d3("category10")(10)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))

# Three 20 color palettes exist for D3
show_col(pal_d3("category20")(20))
show_col(pal_d3("category20b")(20))
show_col(pal_d3("category20c")(20))
```

```{r}
# LocusZoom plaette
# Original Colors
original <- pal_locuszoom()(7)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))
```

```{r}
# Integrated Genomics Viewer Palette
# Original palette has 51 colors
# Variable name is used in title output by swatch() 
`IGV Palette` <- pal_igv()(51)
swatch(`IGV Palette`)

# 20-color subset
original <- pal_igv()(20)
show_col(original)
```

```{r}
# University of Chicago Palette
# Three palettes provided in pal_uchicago
for (pal_opt in c("default", "light", "dark")){
  # Original Colors
  original <- pal_uchicago(pal_opt)(9)
  show_col(original)

  # Extended Colors
  show_col(colorRampPalette(original)(20))
}
```

```{r}
# Star Trek Palette
# Original Colors
original <- pal_startrek()(7)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))
```

```{r}
# Rick and Morty Palette
# Original Colors
original <- pal_rickandmorty("schwifty")(12)
show_col(original)

# Extended Colors
show_col(colorRampPalette(original)(20))
```

```{r}
# University of Colorado Palette
# A palette created in-house using the official colors of CU
# (https://www.colorado.edu/brand/how-use/color)
colors <- c(
  "#000000", 
  "#565A5C", 
  "#A2A4A3",
  "#CFB87C") 
# Original colors
show_col(colors)

# Extended colors
show_col(colorRampPalette(colors)(20))
# This makes a nice continuous palette

# Additional colors in CU Palette
cu_extended <-
  colors <- c(
    # First four: official CU Colors
    "#CFB87C", 
    "#000000", 
    "#565A5C", 
    "#A2A4A3",
    # Color of walls on most campus buildings 
    "#AD7B64",
    # Color of many roofs on campus
    "#9A4A3A",
    # Color of pine trees in Flatirons
    "#3A553A",
    # Color of the sky on a sunny day
    "#529FDF"
    )

show_col(cu_extended)

# Extend to 20 colors
show_col(colorRampPalette(cu_extended)(20))
```

## Continuous Palettes

### ViridisLite

```{r}
show_col(c("black", "#DF536B", "#61D04F", "#2297E6", "#28E2E5", "#CD0BBC", "#F5C710", "gray62"))
```

